---
title: "Hierarchical Spatial Interpolation of Extremes"
author: "Siegfred Roi L. Codia"
subtitle: Prerequisite to MS Stat Thesis
output:
  word_document:
    toc: yes
  html_document:
    toc: yes
    toc_float: yes
    df_print: paged
---

## Importing Required Data

### PH Provincial boundary data

Data obtained from NAMRIA

```{r message=FALSE}
library(geojsonio)
library(broom)
provinces <- geojson_read("Data/Admin Boundaries/provinces.geojson",  what = "sp")
provinces <- tidy(provinces)
```

```{r message=FALSE}
library(tidyverse)
ph_map <- ggplot() +
    geom_polygon(data = provinces,
                 aes( x = long, y = lat,
                      group = group), fill="gray") +
    theme_void() +
    coord_map()+
    theme_bw()
ph_map
```

### Rainfall data

```{r message=FALSE}
stations <- read_csv("NOAA PH Weather Stations.csv")
max_data <- read_csv("max.spread.fixed.csv")
```

`max_data` contains annual rainfall maxima from each station obtained from NOAA GSOD server.

`stations` data contains coordinates and historical maxima of each stations as posted by PAGASA

```{r}
ph_map +
        geom_point(data = stations,
                   aes(x = LON, y = LAT, color = `Rx1day (mm)`))+
        scale_color_viridis_c()+
    ggtitle("Historical Rainfall Maxima Recorded \nby PAGASA Stations")
```

### Digital Elevation Model data of the Philippines

Data is obtained from [The Humanitarian Data Exchange](https://data.humdata.org/dataset/philippines-elevation-model "Philippines - Elevation Model")

```{r}
library(raster)
grd_file <- raster("Data/PHL_ALT/PHL_msk_alt.grd")
elev_df <- as.data.frame(rasterToPoints(grd_file))
colnames(elev_df) <- c("x", "y", "ALT")
```

```{r}
library(ggplot2)
xlabs <- c(115, 120, 125, 130)
ylabs <- c(5,10,15,20)
ggplot()+
    geom_tile(data = elev_df, aes(x = x, y = y, fill = ALT))+
    scale_fill_gradientn(colors = terrain.colors(5))+
    coord_fixed()+
    theme_bw()+
    xlab("LON") + ylab("LAT")+
    scale_x_continuous(breaks = xlabs, labels = paste0(xlabs,'°W'))+
    scale_y_continuous(breaks = ylabs, labels = paste0(ylabs,'°N'))+
    theme(legend.title = element_blank(),
          legend.key.width = unit(0.3, "cm"),
          legend.key.height = unit(1.8, "cm"))+
    ggtitle("Philippine Elevation in meters")
```

## **Fitting pointwise GEVD distributions**

We first load the EnvStats package and create custom function to extract parameters directly.

```{r message=FALSE}
library(EnvStats)
gevd_pars <- function(x){egevd(x)$parameters}
```

Now performing the estimation per column in the data frame.

```{r warning=FALSE}
parameters <- max_data[, -1]%>%sapply(gevd_pars) %>% t() %>% data.frame()
```

Creating the data frame that contains the GEVD parameters per weather station...

```{r}
parameters$"Station Name" <- row.names(parameters)
stations_pars <- full_join(stations,parameters, by = "Station Name")
stations_pars
```

### Visualization

```{r}

library(tidyverse)
rp <- c("5-year", "50-year", "100-year")
rl_gathered <- stations_pars |> 
    mutate("5-year" = qgevd(1-1/5, location = location,
                            scale = scale, shape = shape),
           "50-year" = qgevd(1-1/50, location = location, 
                            scale = scale, shape = shape),
           "100-year" = qgevd(1-1/100, location = location, 
                            scale = scale, shape = shape)) |> 
    dplyr::select(`Station Name`, LAT, LON, all_of(rp)) |> 
    pivot_longer(cols= rp, 
                 names_to = "Return Period",
                 values_to = "Return Level") |> 
    mutate("Return Period" = ordered(`Return Period`, 
                                     levels = rp))
```

```{r}
ph_map +
    geom_point(data = rl_gathered, aes(x = LON, y = LAT, color = `Return Level`))+
    facet_wrap(.~`Return Period`)+
    scale_color_viridis_c()+
    theme_bw()+
    ggtitle("Rainfall Return Level (mm) in the Philippines per weather station")+

    xlab("LON") + ylab("LAT")+
    scale_x_continuous(breaks = xlabs, labels = paste0(xlabs,'°W'))+
    scale_y_continuous(breaks = ylabs, labels = paste0(ylabs,'°N'))+
    theme(plot.title=element_text(hjust=0.5),
          legend.title = element_blank(),
          legend.key.width = unit(0.3, "cm"),
          legend.key.height = unit(1.5, "cm"))
```

### Relationships of Altitude and the GEV parameters

We also explore the relationship of Altitude and the estimated GEV parameters so we will have a basis if we will include them in the spatial interpolation models.

```{r}
stations_pars |> pivot_longer(cols = c(location, scale, shape),
                              names_to = "parameter") |> 
ggplot(aes(x = ALT, y = value))+
    geom_point()+
    geom_smooth(method = "lm", se = F)+
    facet_wrap(.~parameter)
```

```{r}
cor(stations_pars[,c("ALT","location","scale","shape")])
```

```{r}
cor.test(stations_pars$ALT, stations_pars$location)
```

```{r}
cor.test(stations_pars$ALT, stations_pars$scale)
```

While the location and scale parameters have a significant positive correlation with the altitude, we still decide **not to include altitude in spatial interpolation** of the parameters since the correlation coefficient is not high enough (weak to moderate correlation only). To add, the following linear regression outputs have a very low $R^2$ values.

```{r}
lm(location~ALT, data = stations_pars)%>%summary()
```

```{r}
lm(scale~ALT, data = stations_pars)%>%summary()
```

## **Spatial Interpolation**

Recall that the `stations_pars` data frame contains longitude and latitude coordinates of each weather station, and the corresponding estimated GEVD parameters of the annual maxima. This will be the source data for interpolation.

Predictions will be made on the data frame `elev_df` which contains coordinates of Philippine land masses with cell size $0.0083^{\circ}W \times 0.0083^{\circ}N$.

We will convert the two data frames to SpatialDataFrame for better data manipulation using functions for spatial data analysis.

```{r}
library(sp)
# converting the dataframes to spatial data frames
stations_pars_spdf <- SpatialPointsDataFrame(stations_pars,
                                             coords = stations[,c("LON", "LAT")])
elev_spdf <- SpatialPointsDataFrame(elev_df,
                                             coords = elev_df[,c("x", "y")])
```

For a short demonstration of hierarchical spatial interpolation of extremes, we will perform three spatial interpolation procedures of the GEV parameters $\mu,\sigma,\xi$ .

### **Inverse Distance Weighting**

```{r message=FALSE}
idw_loc   <- gstat::idw(location ~ 1, stations_pars_spdf, elev_spdf, idp = 2.0)
idw_scale <- gstat::idw(scale ~ 1, stations_pars_spdf, elev_spdf, idp = 2.0)
idw_shape <- gstat::idw(shape ~ 1, stations_pars_spdf, elev_spdf, idp = 2.0)
```

```{r}
df_idw <- data.frame(x = elev_df$x, y = elev_df$y,
                     location = idw_loc$var1.pred,
                     scale = idw_scale$var1.pred,
                     shape = idw_shape$var1.pred)
```

Computation of return level values

```{r}
library(EnvStats)

rl_idw <- df_idw |> 
    mutate("5-year" = qgevd(1-1/5, location = location,
                            scale = scale, shape = shape),
           "50-year" = qgevd(1-1/50, location = location, 
                            scale = scale, shape = shape),
           "100-year" = qgevd(1-1/100, location = location, 
                            scale = scale, shape = shape)) |> 
    dplyr::select(x, y, all_of(rp)) |> 
    pivot_longer(cols= rp, 
                 names_to = "Return Period",
                 values_to = "Return Level") |> 
    mutate("Return Period" = ordered(`Return Period`, levels = rp))
rl_idw
```

```{r}
ggplot()+
    geom_tile(data = rl_idw, aes(x = x, y = y, fill = `Return Level`))+
    scale_fill_viridis_c(limits = c(0, 1200), oob = scales::squish)+
    coord_fixed()+
    theme_bw()+
    facet_wrap(.~`Return Period`)+
    ggtitle("Rainfall Return Level (mm) in the Philippines using IDW")+
    
    xlab("LON") + ylab("LAT")+
    scale_x_continuous(breaks = xlabs, labels = paste0(xlabs,'°W'))+
    scale_y_continuous(breaks = ylabs, labels = paste0(ylabs,'°N'))+
    theme(plot.title=element_text(hjust=0.5),
          legend.title = element_blank(),
          legend.key.width = unit(0.3, "cm"),
          legend.key.height = unit(1.5, "cm"))
```

### **Kriging**

We will perform an Ordinary Kriging. To determine which type of variogram will be used, the following checks are performed:

```{r}
library(gstat)
v_loc <- variogram(location ~ 1, stations_pars_spdf)

fit_loc_sph <- fit.variogram(v_loc, vgm(model = "Sph"))
fit_loc_lin <- fit.variogram(v_loc, vgm(model = "Lin"))
fit_loc_pow <- fit.variogram(v_loc, vgm(model = "Pow"))

pred_sph <- gstat::krige(location ~ 1,
                           stations_pars_spdf,
                           stations_pars_spdf,
                           model = fit_loc_sph)

pred_lin <- gstat::krige(location ~ 1,
                           stations_pars_spdf,
                           stations_pars_spdf,
                           model = fit_loc_lin)
pred_pow <- gstat::krige(location ~ 1,
                           stations_pars_spdf,
                           stations_pars_spdf,
                           model = fit_loc_pow)

Metrics::rmse(stations_pars_spdf$location, pred_sph$var1.pred)
Metrics::rmse(stations_pars_spdf$location, pred_lin$var1.pred)
Metrics::rmse(stations_pars_spdf$location, pred_pow$var1.pred)

Metrics::mae(stations_pars_spdf$location, pred_sph$var1.pred)
Metrics::mae(stations_pars_spdf$location, pred_lin$var1.pred)
Metrics::mae(stations_pars_spdf$location, pred_pow$var1.pred)
```

Spherical Model has the lowest RMSE and MAE. Hence, we will use Spherical Variogram Model Type in our Ordinary Kriging Spatial Interpolation of the parameters.

```{r message=FALSE}

v_loc <- variogram(location ~ ALT, stations_pars_spdf)
plot(v_loc)
fit_loc <- fit.variogram(v_loc, vgm(model = "Sph"))
krig_loc   <- gstat::krige(location ~ 1,
                           stations_pars_spdf,
                           elev_spdf,
                           model = fit_loc)

v_scale <- variogram(scale ~ 1, stations_pars_spdf)
fit_scale <- fit.variogram(v_scale, vgm(model = "Sph"))
krig_scale <- gstat::krige(scale ~ 1,
                           stations_pars_spdf,
                           elev_spdf,
                           model = fit_scale)

v_shape <- variogram(shape ~ 1, stations_pars_spdf)
fit_shape <- fit.variogram(v_shape, vgm(model = "Sph"))
krig_shape <- gstat::krige(shape ~ 1,
                           stations_pars_spdf,
                           elev_spdf,
                           model = fit_shape)
```

Creating data frame for interpolated values of the parameters.

```{r}
df_krig <- data.frame(x = elev_df$x, y = elev_df$y,
                     location = krig_loc$var1.pred,
                     scale = krig_scale$var1.pred,
                     shape = krig_shape$var1.pred)
```

Computation of return level values

```{r}
rl_krig <- df_krig |> 
     mutate("5-year" = qgevd(1-1/5, location = location,
                            scale = scale, shape = shape),
           "50-year" = qgevd(1-1/50, location = location, 
                            scale = scale, shape = shape),
           "100-year" = qgevd(1-1/100, location = location, 
                            scale = scale, shape = shape)) |> 
    dplyr::select(x, y, all_of(rp)) |> 
    pivot_longer(cols= rp, 
                 names_to = "Return Period",
                 values_to = "Return Level") |> 
    mutate("Return Period" = ordered(`Return Period`, levels = rp))
rl_krig
```

Plotting return level values

```{r}
ggplot()+
    geom_tile(data = rl_krig, aes(x = x, y = y, fill = `Return Level`))+
    scale_fill_viridis_c(limits = c(0, 1200), oob = scales::squish)+
    coord_fixed()+
    theme_bw()+
    facet_wrap(.~`Return Period`)+
    ggtitle("Rainfall Return Level (mm) in the Philippines using Kriging")+
    
    xlab("LON") + ylab("LAT")+
    scale_x_continuous(breaks = xlabs, labels = paste0(xlabs,'°W'))+
    scale_y_continuous(breaks = ylabs, labels = paste0(ylabs,'°N'))+
    theme(plot.title=element_text(hjust=0.5),
          legend.title = element_blank(),
          legend.key.width = unit(0.3, "cm"),
          legend.key.height = unit(1.5, "cm"))
```

### **Spline**

```{r message=FALSE, warning=FALSE}
library(fields)

spline_loc    <- Tps(coordinates(stations_pars_spdf),
                    stations_pars_spdf$location, lambda = 0)
spline_scale   <- Tps(coordinates(stations_pars_spdf),
                    stations_pars_spdf$scale, lambda = 0)
spline_shape   <- Tps(coordinates(stations_pars_spdf),
                    stations_pars_spdf$shape, lambda = 0)
```

```{r}
df_spline <- elev_df |> 
    mutate(location = predict(spline_loc, coordinates(elev_spdf)),
           scale  = predict(spline_scale, coordinates(elev_spdf)),
           shape = predict(spline_shape, coordinates(elev_spdf)))

```

Computation of return level values

```{r}
rl_spline <- df_spline |> 
    mutate("5-year" = qgevd(1-1/5, location = location,
                            scale = scale, shape = shape),
           "50-year" = qgevd(1-1/50, location = location, 
                            scale = scale, shape = shape),
           "100-year" = qgevd(1-1/100, location = location, 
                            scale = scale, shape = shape)) |> 
    dplyr::select(x, y, all_of(rp)) |> 
    pivot_longer(cols= rp, 
                 names_to = "Return Period",
                 values_to = "Return Level") |> 
    mutate("Return Period" = ordered(`Return Period`, levels = rp))
```

```{r}

ggplot()+
    geom_tile(data = rl_spline, aes(x = x, y = y, fill = `Return Level`))+
    scale_fill_viridis_c(limits = c(0, 1200), oob = scales::squish)+
    coord_fixed()+
    theme_bw()+
    facet_wrap(.~`Return Period`)+
    ggtitle("Rainfall Return Level (mm) in the Philippines using Spline")+
    
    xlab("LON") + ylab("LAT")+
    scale_x_continuous(breaks = xlabs, labels = paste0(xlabs,'°W'))+
    scale_y_continuous(breaks = ylabs, labels = paste0(ylabs,'°N'))+
    theme(plot.title=element_text(hjust=0.5),
          legend.title = element_blank(),
          legend.key.width = unit(0.3, "cm"),
          legend.key.height = unit(1.5, "cm"))
```

In the spline interpolation, from the graphs, there is a tendency that this model will over estimate when it comes to extrapolation of values outside the study region


