---
title: "Simulation of Spatial Data"
author: "Siegfred Codia"
output: html_notebook
---

For this example, we simulate normal data over a 40 x 40 grid. The function `mvnorm` from `MASS` library can sample from a multivariate normal distribution. Note that the simulated data assumes spatial stationarity and isotropy.

## Initialization

In a spatial statistics, Matern Covariance function is commonly used. Matern covariance function specifies covariance between two measurements as a function of distance between the points at which they are taken. The Matern covariance between measurements taken at two points separeted by $d$ distance units is given by

$$
C_\nu(d)=\sigma^2 \frac{2^{(1-\nu)}}{\Gamma(\nu)} \left(\sqrt{2\nu} \frac{d}{\alpha}\right)^\nu K_\nu\left(\sqrt{2\nu} \frac{d}{\alpha}\right)
$$

where

-   $\sigma^2$ is the sill parameter
-   $\alpha >0$ is the scale parameter
-   $\nu>0$ is a regularity (smoothness) parameter which determines the mean-square differentiability of the Gaussian field. It measures number of mean square derivatives and is equivalent to the polynomial tail behavior of the spectral density. When
    -   $\nu$ = 0.5 $\rightarrow$ Matern is an exponential covariance

    -   $\nu=\infty \rightarrow$ Matern is Gaussian covariance
-   $\Gamma$ is the gamma function
-   $K_\nu$ is the modified Bessel function of the second kind and order $\nu$

```{r}
# initial function: for creating covariance matrix
cov_matern <- function(x, nu = 2, alpha = 1, vars=1){
    if(nu == 0.5){
        return(vars*exp( - x * alpha))
    }
    
    ismatrix <- is.matrix(x)
    
    if(ismatrix){
        nr=nrow(x)
        nl=ncol(x)
        }
    
    x <- c(alpha * x)
    output <- rep(1, length(x))
    n <- sum(x > 0)
    
    if(n > 0) {
        x1 <- x[x > 0]
        output[x > 0] <-
            (1/((2^(nu - 1)) * gamma(nu))) * (x1^nu) * besselK(x1, nu)
    }
    if(ismatrix){
        output <- matrix(output, nr, nl)
    }
    vars*output
}
```

### Conversion of coordinates to index

The following is a function that converts coordinates to index from the vector of simulated data:

$$
index = 40(y - 1)+x
$$

```{r}
coord0 <- function(x,y){40 * (y - 1) + x}
```

### 

Loading packages `MASS` for simulation of multivariate normal distribution and `ggplot2` for some visualization.

```{r warning=TRUE}
library(MASS)
library(ggplot2)
```

Initializing the 40 x 40 grid and coordinates.

```{r}
xgrid0 <- (rep(1:40, each = 40)-0.5)/40
ygrid0 <- (rep(1:40, times = 40)-0.5)/40
locs0  <- cbind(x = xgrid0,
               y = ygrid0)
# Pairwise distances of coordinates
locs.dist0 <- as.matrix(dist(locs0))

ggplot(data = as.data.frame(locs0), aes(x = x, y = y))+
    geom_point()+theme_bw()+coord_fixed()
```

## Simulation 1: $\nu=\frac{1}{2}, \alpha=7\sqrt{3}$

```{r}
V1 <- cov_matern(locs.dist0, nu=1/2, alpha=7*sqrt(3))
set.seed(20)
z1 <- MASS::mvrnorm(mu=rep(0, 40^2), Sigma=V1)
z1 <- matrix(z1, ncol=40)
```

```{r}
persp(x = (1:40-0.5)/40, y = (1:40-0.5)/40, z1, theta=-45, phi=35, r=5, expand=0.6, axes=T,
      ticktype="detailed", xlab="x", ylab="y", zlab="z1")
```

```{r}
filled.contour(x=((1:40)-0.5)/40, y=((1:40)-0.5)/40, z1, color.palette=gray.colors)
```

## Simulation 2: $\nu=2, \alpha=7\sqrt{3}$

```{r}
V2 <- cov_matern(as.matrix(dist(locs0)), nu = 2, alpha=7*sqrt(3))
z2 <- MASS::mvrnorm(mu=rep(0, 40^2), Sigma=V2)
z2 <- matrix(z2, ncol=40)
```

```{r}
persp(x=((1:40)-0.5)/40, y=((1:40)-0.5)/40, z2, 
      theta=-45, phi=35, r=5, expand=0.6, axes=T,
      ticktype="detailed", xlab="x", ylab="y", zlab="z2")
```

```{r}
filled.contour(x=((1:40)-0.5)/40, y=((1:40)-0.5)/40, z2, color.palette=gray.colors)
```

## Simulation 3: $\nu=10, \alpha=7\sqrt{3}$

```{r}
V3 <- cov_matern(as.matrix(dist(locs0)),nu=10, alpha =7*sqrt(3))
set.seed(100)
z3 <- MASS::mvrnorm(mu = rep(0, 40^2), Sigma=V3)
z3 <- matrix(z3, ncol=40)
```

```{r}
persp(x=((1:40)-0.5)/40, y=((1:40)-0.5)/40, z3, theta=-45, phi=35, r=5, expand=0.6, 
      axes=T,ticktype="detailed", xlab="x", ylab="y", zlab="z")
```

```{r}
filled.contour(x = ((1:40)-0.5)/40, y = ((1:40)-0.5)/40, z3, color.palette=gray.colors)
```

## Creating a GIF

```{r}
library(lattice)
wireframe(z3, shade = TRUE,distance=0,
        screen=list(z=50,x=-60),
        xlab="",ylab="",zlab="",scales=list(draw=FALSE))
```

```{r}
draw.plot <- function(data,angles)
{
  for(i in 1:length(angles))
  {
    print(wireframe(data, shade = TRUE,distance=0,
        screen=list(z=angles[i],x=-60),
        xlab="",ylab="",zlab="",scales=list(draw=FALSE)))
    setTxtProgressBar(txtProgressBar(style=3),i/length(angles))
  }
}
angles <- seq(from = 1,to = 360, by=1)
```

```{r}
library(animation)
saveGIF(draw.plot(z3,angles), interval = 1/10, 
        movie.name = "z3.GIF", 
        ani.height = 640, ani.width = 640)
```
